---
title: "Reading Data from the Web"
author: "Lectured by Jeff Goldsmith"
date: "2022-10-13"
output: github_document
---

```{r setup, echo = FALSE}
library(tidyverse)
library(rvest)
library(httr)
```

# Extracting Tables

```{r load data from web}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

# I want to pull out all the tables from the html webpage:
drug_use_html %>% html_table()
# Oops...Too many tables.....
```

We’re only focused on the first table for now, so let’s get the contents from the first list element.
```{r I only want to look at the first table from the webpage}
table_marj = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
# Notice that the “note” at the bottom of the table appears in every column in the first row.
# We need to remove that…
  slice(-1) 

table_marj
```


# Learning Assessment

Create a data frame that contains the cost of living table for New York from the below html link.
```{r LA1}
col_df = 
  read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") %>% 
  html_table(header = TRUE) %>% 
  first()

col_df
```


# CSS Selectors

Suppose we’d like to scrape the data about the Star Wars Movies from the IMDB page: I want to create a dataframe that shows all Star Wars movies'
title, time length, and revenue made.

```{r movie title length revenue}
# First step, read the html.
starwars_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```

For each element, I’ll use the CSS selector in html_elements() to extract the relevant HTML code, and convert it to text. Then I can combine these into a data frame.
*Green box is something that we want to select.
*Yellow box is all other things that are under the same CSS tag.
*Red box is things that we don't want to include.
```{r movies title time length and revenue}
movies_title = 
  starwars_html %>%
  # on the webpage, use CSS selector to select all movies' titles.
  html_elements(".lister-item-header a") %>%
  # then convert to text.
  html_text()

movies_revenue = 
  starwars_html %>%
  html_elements(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

movies_runtime = 
  starwars_html %>%
  html_elements(".runtime") %>%
  html_text()

starwars_movies = 
  tibble(
    title = movies_title,
    revenue = movies_revenue,
    runtime = movies_runtime)

starwars_movies
```


# Learning Assessment 2

The webpage (url) below contains the 10 most recent reviews of the movie “Napoleon Dynamite”. Use a process similar to the one above to extract the titles of the reviews.

```{r LA2}
amazon_html = 
  read_html("https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber=1")

review_titles = 
  amazon_html %>% 
  html_elements(".a-text-bold span") %>% 
  html_text()

review_stars = 
  amazon_html %>% 
  html_elements("#cm_cr-review_list .review-rating") %>% 
  html_text()

review_comments = 
  amazon_html %>% 
  html_elements(".review-text-content span") %>% 
  html_text()

movie_reviews = 
  tibble(
    title = review_titles,
    stars = review_stars,
    comments = review_comments)

movie_reviews
```


# Using an API


