---
title: "Reading Data from the Web"
author: "Lectured by "Jeff Goldsmith""
date: "2022-10-13"
output: github_document
---

```{r setup, echo = FALSE}
library(tidyverse)
library(rvest)
library(httr)
```

# Extracting Tables

```{r load data from web}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

drug_use_html %>% html_table()
```

We’re only focused on the first table for now, so let’s get the contents from the first list element.
```{r}
table_marj = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
# Notice that the “note” at the bottom of the table appears in every column in the first row.
# We need to remove that…
  slice(-1) 
```

# CSS Selectors

Suppose we’d like to scrape the data about the Star Wars Movies from the IMDB page. The first step is the same as before – we need to get the HTML.
```{r}
swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```
For each element, I’ll use the CSS selector in html_elements() to extract the relevant HTML code, and convert it to text. Then I can combine these into a data frame.
```{r}
title_vec = 
  swm_html %>%
  html_elements(".lister-item-header a") %>%
  html_text()

gross_rev_vec = 
  swm_html %>%
  html_elements(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

runtime_vec = 
  swm_html %>%
  html_elements(".runtime") %>%
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    rev = gross_rev_vec,
    runtime = runtime_vec)
```

# Using an API


