---
title: "Strings and Factors"
author: "Lectured by Jeff Goldsmith"
date: "2022-10-18"
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(rvest)
library(p8105.datasets)
```

# Strings and regex

You can use `str_detect` to find cases where the match exists (often useful in conjunction with `filter`), 
and you can use `str_replace` to replace an instance of a match with something else (often useful in conjunction with `mutate`). 

In the following examples we’ll mostly focus on vectors to avoid the complication of data frames, but we’ll see those shortly.

```{r EX1}
# Let us define a collection of string vectors, they will be presented in dataframe:
string_vec = c("my", "name", "is", "jeff")

# Detect this pattern "jeff" in the string_vec:
str_detect(string_vec, "jeff")

# Detect "a" in the string_vec:
str_detect(string_vec, "a")
# "a" is only appeared in the second string "name".

# Detect "m":
str_detect(string_vec, "m")
# "m" is appeared in the first two strings "my" & "name".
```

```{r EX2}
# Replace the string "jeff" with "Jeff":
str_replace(string_vec, "jeff", "Jeff")

# Replace lowercase "m" with uppercase "M":
str_replace(string_vec, "m", "M")
```

For exact matches, we can designate matches at the beginning or end of a line.
```{r match strings}
string_vec2 = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think")

str_detect(string_vec2, "i think")
# It will return "TRUE" for all because "i think" appears in all four strings.

# Detect something that starts with "i think":
str_detect(string_vec2, "^i think")

# Detect something that ends with "i think":
str_detect(string_vec2, "i think$")
```

In the following example, we see "Bush" in different forms (upper/lowercase, standalone word or connected with other words). We can still detect all of them:
```{r EX3}
string_vec3 = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!")

str_detect(string_vec3,"[Bb]ush")
# The all uppercase BUSH is not detected.

# This will allow me to detect all forms of "bush".
str_detect(string_vec3, "[Bb][Uu][Ss][Hh]")

```

Which of these following vectors have numbers followed by letters?
```{r EX4}
string_vec4 = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :(')

# Detect anything that has a number from 0 to 9, followed by a letter from capital A to Z:
str_detect(string_vec4, "[0-9][A-Z]")

# Detect anything that has a number from 0 to 9, followed by a letter from lowercase a to z or capital A to Z:
str_detect(string_vec4, "[0-9][a-zA-Z]")
```

The character `.` matches anything.
```{r EX5}
string_vec5 = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66')

str_detect(string_vec5, "7.11")
# The third string wasn't detected because there isn't anything between 7 & 11.
```

What if I want to search for something exactly like numbers in the []? Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. 
If we want to search for these, we have to indicate they’re special using `\`. Unfortunately, `\` is also special, so things get weird.
```{r EX6}
string_vec6 = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]')

# str_detect(string_vec6, "[")
# This is not going to work because R thinks you forgot to close the bracket.

# If we want to look for exactly this "[]" in the strings, we need a "\" to indicate the bracket is special. 
# But because the "\" is also a special character, we need two "\" to make this work:
str_detect(string_vec6, "\\[")

# Detect something that has a bracket "[" followed by a range of numbers 0 to 9:
str_detect(string_vec6, "\\[[0-9]")
```

# Why factors are weird

```{r EX7}
factor_sex = factor(c("male", "male", "female", "female"))
# R tells us there are 2 levels of factors: female male. The order of female is ahead of male, so female is 1 and male is 2.

# If I make the factor vector numeric, it will give me 2 2 1 1:
as.numeric(factor_sex)

# I can relevel my factor vector:
factor_sex = fct_relevel(factor_sex, "male") # I am telling R to make the order of "male" before "female".

# Now if I do as.numeric I should see 1 1 2 2.
as.numeric(factor_sex)
```


# Let's do some practice!

## *NSDUH*

```{r practice last lecture}
NSDUH = read_html("http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm") %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

There are a few steps we need to implement to tidy these data. For now I’m not interested in the p-values (I’d rather just see the data); we also have age groups and year ranges in the same column titles, 
both of which are, in fact, variables. Lastly, the table includes letters as superscripts next to table entries; if we only want the percents we’ll need to strip these out.
```{r tidy up NSDUH}
NSDUH_marj = 
  NSDUH %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
# If we right away change "percent" into numeric variable, the rows that contain numbers followed by a letter will be forced to become "NA" because R don't know what to do with them.
# So we want to get rid of the letters first before we change the variable to numeric. 
  mutate(
# I want to detect any strings that has letters at the end which range from a to c, and replace them with nothing.
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>% 
# We want to separate "age" "year" by the opened parentheses ( but ( are special character, so we need "\" to indicate that.
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
# But now we have a closed parentheses ) hanging in the end of the strings for year, and we need to remove that.
  mutate(
# Replace the ")" with nothing
    year = str_replace(year, "\\)", "")) %>%
# I just want to look at each 50 states, not Total U.S., Northeast, etc.
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West", "District of Columbia")))
```

Let’s quickly visualize these data for the *12-17* age group; to make the plot readable, we’ll treat `State` as a factor and are reordered according to the median `percent` value.
```{r calculate median and plot}
NSDUH_marj %>%
  filter(age == "12-17") %>% 
# I want to reorder different states by ascending order in "percent" variable.
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
  labs(
    title = "Percentage of marijuana use among 12-17 age group") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# vjust adjust the position of the axis. If I use vjust = 2, my x-axis will move 2 clicks to the right.
# hjust adjust the up-down position of the axis. If I do hjust = 2, my x-axis will move 2 clicks up and overlap with my plot.
```


## *Restaurant Inspection*

The `Restaurant Inspection` dataset is inside the `p8105.datasets` package.

```{r load dataset}
data("rest_inspec")
# There is about 400,000 rows in this dataset
rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(
    names_from = grade, 
    values_from = n)
```

To simplify things, I’ll remove inspections with scores other than A, B, or C, and also remove the restaurants with missing `boro` information. I’ll also clean up `boro` names a bit.
```{r cleaning up a bit}
nyc_rest_insp = 
  rest_inspec %>% 
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro)) # Instead of making all letters lowercase, str_to_title keep the first letter uppercase.

nyc_rest_insp %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(
    names_from = grade, 
    values_from = n)
```

Let’s find how many pizza places are there in NYC.
```{r pizza time}
nyc_rest_insp %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro) %>% 
  count() # summarize(n = n()) will give the same results
  
nyc_rest_insp %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = str_replace(boro, "Manhattan", "The City"), # str_replace allow me to change "Manhattan" into "The City".
    # Can also do: boro = fct_recode(boro, "The City" = "Manhattan")
    boro = fct_infreq(boro)) %>% # I can rearrange the boro in terms of highest frequency to lowest frequency.)  
  ggplot(aes(x = boro)) +
  geom_bar()
```

What about examining pizza places and their grades by borough?
```{r pizza grades}
nyc_rest_insp %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

nyc_rest_insp %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```


## *Weather Data*

```{r load weather data, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Our first example reordered name “by hand” using `fct_relevel`:
```{r hand reorder}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

We could instead reorder name according to tmax values in each name using `fct_reorder`:
```{r reorder by tmax}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

#### **Linear regression**

The ordering of factor variables is important in linear regression. Specifically, the ordering determines the “reference” category, and is something that can be adjusted as needed.
```{r Central Park as reference}
weather_df %>%
  lm(tmax ~ name, data = .) # Since we are piping, "data = . " tells R the tibble it should be using is the one that just piped in.
```

```{r Waikiki as reference}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  lm(tmax ~ name, data = .)
```


## *PULSE Data*

Using `stringr` and `forcats`, the result is the same as we did many lectures before, and the differences are pretty small, but this is a bit cleaner.

```{r stringr forcats, eval = FALSE}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)

print(pulse_data, n = 12)
```


## **Airbnb**

We could also use factors in an exploratory analysis of the Airbnb data, for example when looking at the distribution of prices in various neighborhoods. 
Ordering these according to the median price makes for clearer plots than ordering neighborhoods alphabetically.

```{r airbnb}
data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```


