---
title: "Strings and Factors"
author: "Lectured by Jeff Goldsmith"
date: "2022-10-18"
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(rvest)
library(p8105.datasets)
```

# Strings and regex

You can use `str_detect` to find cases where the match exists (often useful in conjunction with `filter`), 
and you can use `str_replace` to replace an instance of a match with something else (often useful in conjunction with `mutate`). 

In the following examples we’ll mostly focus on vectors to avoid the complication of data frames, but we’ll see those shortly.

```{r EX1}
# Let us define a collection of string vectors, they will be presented in dataframe:
string_vec = c("my", "name", "is", "jeff")

# Detect this pattern "jeff" in the string_vec:
str_detect(string_vec, "jeff")

# Detect "a" in the string_vec:
str_detect(string_vec, "a")
# "a" is only appeared in the second string "name".

# Detect "m":
str_detect(string_vec, "m")
# "m" is appeared in the first two strings "my" & "name".
```

```{r EX2}
# Replace the string "jeff" with "Jeff":
str_replace(string_vec, "jeff", "Jeff")

# Replace lowercase "m" with uppercase "M":
str_replace(string_vec, "m", "M")
```

For exact matches, we can designate matches at the beginning or end of a line.
```{r match strings}
string_vec2 = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think")

str_detect(string_vec2, "i think")
# It will return "TRUE" for all because "i think" appears in all four strings.

# Detect something that starts with "i think":
str_detect(string_vec2, "^i think")

# Detect something that ends with "i think":
str_detect(string_vec2, "i think$")
```

In the following example, we see "Bush" in different forms (upper/lowercase, standalone word or connected with other words). We can still detect all of them:
```{r EX3}
string_vec3 = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!")

str_detect(string_vec3,"[Bb]ush")
# The all uppercase BUSH is not detected.

# This will allow me to detect all forms of "bush".
str_detect(string_vec3, "[Bb][Uu][Ss][Hh]")

```

Which of these following vectors have numbers followed by letters?
```{r EX4}
string_vec4 = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :(')

# Detect anything that has a number from 0 to 9, followed by a letter from capital A to Z:
str_detect(string_vec4, "[0-9][A-Z]")

# Detect anything that has a number from 0 to 9, followed by a letter from lowercase a to z or capital A to Z:
str_detect(string_vec4, "[0-9][a-zA-Z]")
```

The character `.` matches anything.
```{r EX5}
string_vec5 = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66')

str_detect(string_vec5, "7.11")
# The third string wasn't detected because there isn't anything between 7 & 11.
```

What if I want to search for something exactly like numbers in the []? Some characters are “special”. These include `[` and `]`, `(` and `)`, and `.`. 
If we want to search for these, we have to indicate they’re special using `\`. Unfortunately, `\` is also special, so things get weird.
```{r EX6}
string_vec6 = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]')

# str_detect(string_vec6, "[")
# This is not going to work because R thinks you forgot to close the bracket.

# If we want to look for exactly this "[]" in the strings, we need a "\" to indicate the bracket is special. 
# But because the "\" is also a special character, we need two "\" to make this work:
str_detect(string_vec6, "\\[")

# Detect something that has a bracket "[" followed by a range of numbers 0 to 9:
str_detect(string_vec6, "\\[[0-9]")
```

# Why factors are weird

```{r EX7}
factor_sex = factor(c("male", "male", "female", "female"))
# R tells us there are 2 levels of factors: female male. The order of female is ahead of male, so female is 1 and male is 2.

# If I make the factor vector numeric, it will give me 2 2 1 1:
as.numeric(factor_sex)

# I can relevel my factor vector:
factor_sex = fct_relevel(factor_sex, "male") # I am telling R to make the order of "male" before "female".

# Now if I do as.numeric I should see 1 1 2 2.
as.numeric(factor_sex)
```


# NSDUH

```{r practice last lecture}
NSDUH = read_html("http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm") %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

There are a few steps we need to implement to tidy these data. For now I’m not interested in the p-values (I’d rather just see the data); we also have age groups and year ranges in the same column titles, 
both of which are, in fact, variables. Lastly, the table includes letters as superscripts next to table entries; if we only want the percents we’ll need to strip these out.
```{r tidy up NSDUH}
NSDUH_marj = 
  NSDUH %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
# If we right away change "percent" into numeric variable, the rows that contain numbers followed by a letter will be forced to become "NA" because R don't know what to do with them.
# So we want to get rid of the letters first before we change the variable to numeric. 
  mutate(
# I want to detect any strings that has letters at the end which range from a to c, and replace them with nothing.
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>% 
# We want to separate "age" "year" by the opened parentheses ( but ( are special character, so we need "\" to indicate that.
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
# But now we have a closed parentheses ) hanging in the end of the strings for year, and we need to remove that.
  mutate(
# Replace the ")" with nothing
    year = str_replace(year, "\\)", "")) %>%
# I just want to look at each 50 states, not Total U.S., Northeast, etc.
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West", "District of Columbia")))
```

Let’s quickly visualize these data for the *12-17* age group; to make the plot readable, we’ll treat `State` as a factor and are reordered according to the median `percent` value.
```{r calculate median and plot}
NSDUH_marj %>%
  filter(age == "12-17") %>% 
# I want to reorder different states by ascending order in "percent" variable.
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
  labs(
    title = "Percentage of marijuana use among 12-17 age group") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
# vjust adjust the position of the axis. If I use vjust = 2, my x-axis will move 2 clicks to the right.
# hjust adjust the up-down position of the axis. If I do hjust = 2, my x-axis will move 2 clicks up and overlap with my plot.
```





